---
title: "FlowerPetalColorClustering"
author: "Callin Switzer"
date: "September 24, 2018"
output: html_document
editor_options: 
  chunk_output_type: console
---
<pre>
Goal: group images of petals, according to their color
</pre>


# Set options for knitr
```{r setup}
knitr::opts_chunk$set(echo = TRUE)

# define data and figure directories
windows = ifelse(grepl(pattern = "wind", ignore.case = TRUE, 
                       x = Sys.info()[1]), TRUE, FALSE)
  
{if(windows){
  rd = "D:/Dropbox/dataAnalysisForOthers/FlowerColorPhenotyping/"
}
else{
  rd = "/Users/cswitzer/Dropbox/dataAnalysisForOthers/FlowerColorPhenotyping/"
}}

knitr::opts_knit$set(root.dir =rd )

```

# Install packages
```{r, warning=FALSE, message = FALSE}
# Setup
# install.packages("imager")
library(imager)
library(ggplot2)
library(tidyverse)
```


# Read in images from a directory
* These images are in the form of png files
* R reads them in with 4 color channels - red, green, blue, and transparency
```{r}
imageDirectory <- "D:\\Dropbox\\dataAnalysisForOthers\\FlowerColorPhenotyping"

# keep only images that end with .png
imageFileNames <- dir(imageDirectory, 
                      full.names = TRUE)[grepl(pattern = "png", 
                                               dir(imageDirectory, 
                                                   full.names = FALSE))]
```

# Clean images (remove transparency)
```{r}
# function to remove the transparency channel
removeAlpha <- function(img){
  # note that this won't work for images with
  # only 3 channels
  
  as.tbl(as.data.frame(img)) %>%
    spread(key = cc, value = value, fill = 0) %>%
    filter(`4` == 1) %>%
    select(-`4`) %>%
    gather(key = cc, value = value, `1`:`3`) %>%
    mutate(cc = as.integer(cc)) 
}


# remove alpha channel for each image
imageList = imlist()
for(ii in 1:length(imageFileNames)){
  tmp = load.image(imageFileNames[ii])
  imageList[[ii]] <- removeAlpha(tmp)
}
```

# Show images  
* Previously transparent parts are shown in black
```{r, warning=FALSE}
par(mfrow = c(3,3))
for(ii in 1:length(imageList)){
  plot(as.cimg(as.data.frame(imageList[[ii]])))
}
```


# Get color information for images
* Make a data frame where each row is in individual flower
```{r, warning = FALSE}
histList = list()
for(jj in 1:length(imageList)){
  
  # convert to colorspaces from RGB to Lab (a perceptually uniform colorspace -
  # the numerical change corresponds to visually perceived change)
  # see: https://en.wikipedia.org/wiki/CIELAB_color_space
  im2 <- RGBtoLab(as.cimg(as.data.frame(imageList[[jj]])))
  
  
  # filter out data where luminance is > 0 (i.e. 
  # get rid of black)
  bdf <- as.tbl(as.data.frame(im2)) %>%
    # remove black
    spread(key = cc, value = value) %>%
    filter(`1` > 0) %>%
    gather(key = cc, value = value, `1`:`3`) 
  
  
  # concatenate histograms for each of L, a, and b
  # into a single vector, so that we can later calculate 
  # a distance matrix
  aa <- c(hist(bdf$value[bdf$cc == "1"], 
               breaks = seq(-100, 100,length.out = 100 ), plot = FALSE)$density, 
          hist(bdf$value[bdf$cc == "3"], 
               breaks = seq(-100, 100,length.out = 100 ), plot = FALSE)$density, 
          hist(bdf$value[bdf$cc == "3"], 
               breaks = seq(-100, 100,length.out = 100 ), plot = FALSE)$density)
  # append to list
  histList[[jj]] = aa
  
}

# show example of historgrams in Lab colorspace
# ggplot(bdf,aes(value,col=cc)) + 
#   geom_histogram(bins=30)+
#   facet_wrap(~ cc)


# convert list to tbl
labDF <- as.tbl(as.data.frame(do.call(rbind, histList)))


# drop columns that are duplicated
labDF <- labDF[!duplicated(as.list(labDF))]

# double check that all rows have been standardized
rowSums(labDF)
```

# Cluster images using heirarchical clustering
* Note that many other clustering methods are available, such as K-nearest neighbors
```{r, warning = FALSE}
# first calculate distance matrix
# dist(): compute the distances between the rows of a data matrix
distMatrix <- dist(labDF)

# use heirarchical clustering
hcl = hclust(distMatrix)

# plot dendrogram and images
pdf("ClusterAndImages.pdf")
par(mar = c(2,2,2,2))
m <- rbind(c(1, 1,1,1,1),c(1, 1,1,1,1), c(2, 3,4,5,6), c(7,8,9,10,11))
layout(m)

#plot dendrogram
plot(hcl)

# plot images
for(ii in 1:length(imageList)){
  plot(as.cimg(as.data.frame(imageList[[ii]])), main = ii, ylim = c(0, 500))
}
dev.off()

```
